Predictive 3D modelling of free oblique cutting introducing an ANN-based material flow law with experimental validation over a wide range of conditions

### 2023-12-12 Article.tex
Update paper with some adds in red

The material model of Ti6Al4V used in all the numerical simulations proposed in the section \ref{sec:ExpNumResults} is a thermo-elasto-viscoplastic law using a flow criterion based on an Artificial Neural Network (ANN) identified for the chosen material and implemented in the Abaqus/Explicit code via a Fortran subroutine VUHARD as proposed by Pantalé et al. in \cite{pantale_Efficient_2022}.
The principle of this approach consists in replacing the analytical formulation of the flow law, based on a Johnson-Cook or Zerilli-Armstrong type model, and allowing the calculation of the flow stress $\sigma^y$ as a function of the plastic strain $\varepsilon^p$, the plastic strain rate, $\mdot{\varepsilon}^p$, and the temperature $T$, by a multi-layer ANN serving as universal approximator. Thus, the parameters of the neural network can be directly identified from the experimental data without having to postulate a behavioural model, which simplifies the procedure and allows greater flexibility in the definition of the model.
The proposed approach also allows, as shown in Pantalé et al. \cite{pantale_Efficient_2022}, to compute the three derivatives of the flow stress $\sigma^y$ with respect to the three input variables of the model, a necessary step to implement this model as a flow law in the form of a VUHARD subroutine in the FEM code Abaqus/Explicit, using the exact same network architecture and identified trained parameters as the one used to compute the flow stress $\sigma^y$.

In order to verify the influence of the neural network complexity on the numerical results of the simulation and on the computation time, several ANN architectures are tested afterwards (in \ref{subsec:nberneu}). The chosen global architecture has 2 hidden layers with a variable number of neurons for the first hidden layer ($\zeta$~=~9 to 17) and 7 neurons for the second hidden layer, 3 inputs (the plastic strain, $\varepsilon^p$, the plastic strain rate, $\mdot{\varepsilon}^p$, and the temperature, $T$) and one output (the yield strength, $\sigma^y$). The global architecture of this type of ANN is given in Figure \ref{fig:ANN} for 9 neurons in the first hidden layer. According to Pantalé et al. \cite{pantale_Efficient_2022}, this ANN is referred to as ANN 3-9-7-1-sig, as it has 3 inputs, 9 neurons in the first hidden layer, 7 neurons in the second hidden layer, 1 output and a sigmoid activation function.

@@

In a preliminary phase, after having selected the global architecture of the neural network, it is necessary to proceed to its training from some inputs.
The inputs for this application were generated from the Johnson-Cook flow law expression reported in Equation (\ref{eq:J-C}) and the identified parameters reported in Table \ref{tab:prop}. This approach was chosen to demonstrate the ability of the neural network flow law to replace a classically formulated flow law such as Johnson-Cook's for the simulation of metal cutting. In future developments, experimental tests on a Gleeble thermomechanical simulator will be used to generate this network training data. The training data, presented in the form of a data table containing the plastic strain $\varepsilon^p$, the plastic strain rate $\mdot{\varepsilon}^p$, the temperature $T$ and the flow stress $\sigma^y$, is processed by a learning algorithm, developed at LGP, in Python, using the Tensorflow library. One hour of training on a Dell XPS13 7390 laptop running Ubuntu 20.04 64~bits with 16~GiB of Ram and an Intel 4-core i7-10510U processor allow obtaining the converged parameters of the ANN model.

>>>>

\textcolor{red}{In the numerical simulations presented in Section \ref{sec:ExpNumResults}, a thermo-elasto-viscoplastic material model for Ti6Al4V is employed, which utilizes a flow criterion based on an Artificial Neural Network (ANN) identified for the material. This ANN is implemented in the Abaqus/Explicit code through a Fortran VUHARD subroutine, as proposed by Pantalé et al. \cite{pantale_Efficient_2022, pantale_Development_2023}, to compute the flow stress $\sigma^y$ as a function of the plastic strain $\varepsilon^p$, the plastic strain rate, $\mdot{\varepsilon}^p$, and the temperature $T$. The approach replaces the analytical formulation of the flow law, typically based on Johnson-Cook or Zerilli-Armstrong type models, with a multi-layer ANN serving as a universal approximator. This enables the direct identification of the neural network parameters from experimental data without postulating a behavioral model, simplifying the procedure and providing greater flexibility in model definition.}

\textcolor{red}{In contrast to the classic approach, which involves conducting experiments on a material, postulating an analytical form for the flow law, and identifying the parameters that best fit the experimental data, the use of ANN allows for direct identification of the law from experimental data without the need to postulate the analytical form of the flow law. This method also enables the computation of the three derivatives of the flow stress $\sigma^y$ with respect to the three input variables of the model, which is necessary for implementing the model as a flow law in the form of a VUHARD subroutine in the FEM code Abaqus/Explicit. The same network architecture and identified trained parameters are used to compute the flow stress $\sigma^y$ and the derivatives in a one-step procedure \cite{pantale_Efficient_2022, pantale_Development_2023}.}

In order to verify the influence of the neural network complexity on the numerical results of the simulation and on the computation time, several ANN architectures \textcolor{red}{(i.e. hyperparameters of the ANN)} are tested afterwards (in \ref{subsec:nberneu}). The chosen global architecture has 2 hidden layers with a variable number of neurons for the first hidden layer ($\zeta$~=~9 to 17) and 7 neurons for the second hidden layer, 3 inputs (the plastic strain, $\varepsilon^p$, the plastic strain rate, $\mdot{\varepsilon}^p$, and the temperature, $T$) and one output (the yield strength, $\sigma^y$). The global architecture of this type of ANN is given in Figure \ref{fig:ANN} for 9 neurons in the first hidden layer. According to Pantalé et al. \cite{pantale_Efficient_2022}, this ANN is referred to as ANN 3-9-7-1-sig, as it has 3 inputs, 9 neurons in the first hidden layer, 7 neurons in the second hidden layer, 1 output and a sigmoid activation function.
\textcolor{red}{The selection of an architecture with two hidden layers was made based on the conclusions drawn in Pantalé et al. \cite{pantale_Efficient_2022}. Additionally, the decision to use the sigmoid activation function was guided by the findings in Pantalé \cite{pantale_Comparing_2023}, who identified the most efficient and accurate activation functions for finite element simulations in thermomechanical forming.}

@@

In a preliminary phase, after having selected the global architecture of the neural network, it is necessary to proceed to its training from some inputs.
The inputs for this application were generated from the Johnson-Cook flow law expression reported in Equation (\ref{eq:J-C}) and the identified parameters reported in Table \ref{tab:prop}. This approach was chosen to demonstrate the ability of the neural network flow law to replace a classically formulated flow law such as Johnson-Cook's for the simulation of metal cutting. In future developments, experimental tests on a Gleeble thermomechanical simulator will be used to generate this network training data. The training data, presented in the form of a data table containing the plastic strain $\varepsilon^p$, the plastic strain rate $\mdot{\varepsilon}^p$, the temperature $T$ and the flow stress $\sigma^y$, is processed by a learning algorithm, developed at LGP, in Python, using the Tensorflow library \cite{Tensorflow-2015}. One hour of training on a Dell XPS13 7390 laptop running Ubuntu 20.04 64~bits with 16~GiB of Ram and an Intel 4-core i7-10510U processor allow obtaining the converged parameters of the ANN model.
